Maximizing Council Participation: A Study on Toronto City Council Absences (2022-2026)

### 2024-09-27 paper/paper.qmd
final small writing changes

The Canadian parliamentary system is a democracy, where the population is represented in government by a selection of elected officials. To maintain the democratic structure, individuals should engage in political participation to have their voices represented, primarily through voting. Likewise, elected officials should uphold their duty of representing the opinions of those who voted them in. Otherwise, failing to do so will cause a decrease in trust in the government. So, it is the job of the elected members of the Toronto City Council to attend their meetings and express the concerns of the population they represent.

@@

The paper is structured as follows: [@sec-data] discusses the raw and clean data used in the study and any limitations in the dataset. [@sec-results] shows a graphical visualization of the data from [@sec-data] and explains some trends that can be seen. Finally, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study.

@@

After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results]. To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr), randomNames (@randomNames) and ggplot2 (@ggplot2). ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.

# generate first 10 observations of cleaned data
kable(head(raw_data, 10)) 
```
The key variables of interest are name, committee, session year, session type, and present. First and last names of City Council members are combined into 1 variable to make it easier to read in graphs. Then, the session date is decoded to separate the year, month and date - keeping only the year. When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset. The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data. Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.

@@

# References

>>>>

The Canadian parliamentary system is a democracy, where the population is represented in government by a selection of elected officials. To maintain the democratic structure, individuals should engage in political participation to have their voices represented, primarily through voting. Likewise, elected officials should uphold their duty of representing the opinions of those who voted them in. Otherwise, failing to do so will cause a decrease in trust in the government. There are 3 levels of governments in Canada: federal, provincial, and municipal. The municipal government works on a city-wide scale, directly working to address local concerns. So, it is the job of the elected members of the Toronto City Council to attend their meetings and express the concerns of the population they represent.

@@

The paper is structured as follows: [@sec-data] discusses the raw and clean data used in the study and any limitations in the dataset. [@sec-results] shows a graphical visualization of the data from [@sec-data] and explains some trends that can be seen. Next, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study. Finally, [@sec-appendix] includes additional supporting data and [@sec-references] contains the references used in this paper.

@@

The data was into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results]. The key variables of interest are name, committee, session year, session type, and present. @tbl-cleaned-data shows first 10 observations of the cleaned data and [@sec-appendix] explains in more detail how the data was cleaned. The following packages in R (@citeR) were used in the data analysis process: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr), randomNames (@randomNames) and ggplot2 (@ggplot2). ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.

@@

# Appendix {#sec-appendix}

# generate first 10 observations of cleaned data
kable(head(raw_data, 10)) 
```

First and last names of City Council members are combined into 1 variable to make it easier to read in graphs. Then, the session date is decoded to separate the year, month and date - keeping only the year. When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset. The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data. Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.

# References {#sec-references}
